---
title: "MTH6139 Time Series"
subtitle: "Coursework 1"
author:
- name: Mohammed Ahad
date: "March 2024"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r, echo=FALSE}
htmltools::img(src = knitr::image_uri("QMlogo.png"), 
               alt = 'logo', 
               style = 'position:absolute; top:0; right:0; padding:10px; width:30%;')
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The Data I will be using is from the ASTSA library and measures the Quarterly interest rate from 1953 to 1980. It is labelled qintr.

### 2.
```{r}
library(astsa)
qintr.dataframe = data.frame(ds=zoo::as.yearmon(time(qintr)), y=qintr)
```
The time series is setup as a data frame called "qintr.dataframe" where "ds" is the time vector and "y" is the data vector. "ds" will be the year and "y" will be the interest rate. 


### 3.
```{r}
library(prophet)
ModelwithnoSeasonality = prophet(qintr.dataframe)
ModelwithSeasonality = prophet(qintr.dataframe , yearly.seasonality=TRUE)
```
I have used the prophet function to analyse the Quarterly interest rate and to create 2 models which use its knowledge of the 
time series data to predict and forecast future values. One model Without seasonality and one model with.
This is important because being able to accurately predict future interest rates could give an investor a major advantage in the financial markets. 

### 4. 
```{r}
forecastwithnoSeasonality = make_future_dataframe(ModelwithnoSeasonality, periods=8, freq="quarter")
forecastwithSeasonality = make_future_dataframe(ModelwithSeasonality, periods=8, freq="quarter")
```
I have created 2 different data frames with 8 future quarters for predictions


### 5. 
```{r}
PredictionwithnoSeasonality =predict(ModelwithnoSeasonality,forecastwithnoSeasonality) 
PredictionwithSeasonality = predict(ModelwithSeasonality,forecastwithSeasonality)
```
Using the earlier prophet model the function will make 8 quarterly predictions, one without seasonality and one with.


### 6.
```{r}
plot(ModelwithnoSeasonality,PredictionwithnoSeasonality)
plot(ModelwithSeasonality,PredictionwithSeasonality)
```

The shaded region represent the prediction interval and the black dots are the observed data point. Both graphs look very similar which shows that there is a minimal seasonal effect . As shown in both graphs the vast majority of the data points lie within the prediction region apart from a few outlier on the tail end. This shows that there is a high degree a accuracy with the meta prophet function.